---
title: "King County House Price"
author: "Yuhsiang Hong"
date: "4/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective

The purpose of this project is to build a regression model that can predict the house price based on some features of the house.

```{r}
library(tidyverse)
library(ggmap)
```

## Data Cleaning

```{r}
# Import the data
kc_house <- read_csv("kc_house_data.csv")
head(kc_house)
summary(kc_house)
```

Since "id" is meaningless to house price, it will be removed. There are several variables that can be modified or merged to more meaningful new variables:

1. "age" will be computed by "date" minus "yr_built" to represent how many years that the house had been built until the year it was sold.

2. There are a lot of houses that have never been renovated, so "yr_renovated" will be converted to a binary variable, "renovated", showing whether this house had been renovated before it was sold.

3. There are a lot of houses without basement as we can see from the summary. Thus, I will modify the "sqft_basement" to a binary variable, "basement", indicating if the house has the basement.

4. Since all the data are from a county, the changes of longtitude and lattitude are very small. Moreover, the dataset contains zipcode, which is sufficient to reflect the relationship between the house price and the location. Therefore, I will remove these two variables.

For zipcode, waterfront and view, I will change them into factorial variables.

```{r}
kc_house <- kc_house %>% 
  mutate(age = as.numeric(format(date, format="%Y")) - yr_built) %>% 
  mutate(renovated = as.factor(as.numeric(yr_renovated != 0))) %>% 
  mutate(basement = as.factor(as.numeric(sqft_basement != 0))) %>% 
  mutate(zipcode = as.factor(zipcode)) %>% 
  mutate(waterfront = as.factor(waterfront)) %>% 
  mutate(view = as.factor(view)) %>% 
  select(-id, -date, -sqft_basement, -yr_built, -yr_renovated)
```

Handle missing values.

```{r}
sapply(kc_house, function(x) sum(is.na(x)))
```

Since there are only two missing values, I will remove these two observations with missing values.

```{r}
kc_house <- na.omit(kc_house)
```

## Data Exploratory

```{r}

```

```{r}
qmplot(long, lat, maptype = "watercolor", color = log(price),
  data = kc_house) + guides(color = FALSE) +
  scale_colour_viridis_c()
```



